\textit{Supplemental materials for:} High-order cognition is supported by complex but compressible brain activity patterns

### 2023-02-27 paper/main.tex
finished draft of methods and supplemental materials

\caption{\textbf{Top terms associated with the highest-weighted components by
condition.} Each row corresponds to an experimental condition, and the colors
correspond to the component number (ranked by proportion of variance
explained). The inflated brain plots display the spatial distributions of each
components' hubs (see \textit{Topographic Factor Analysis}). The lists on the
right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
each components' brain map. Analogous maps computed separately for each story
segment may be found in Figure~\synthThirds.}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/pca_changes_neurosynth}

\caption{\textbf{Changes in decoding accuracy and compression over time.} Each
row corresponds to an experimental condition, and the colors correspond to
change direction (positive: red; negative: blue). The saturation denotes the
component number (columns on the right).  The lists on the
right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
each components' brain map.}

\label{fig:neurosynth-changes}

@@

>>>>

\caption{\textbf{Top terms associated with the highest-weighted components by
condition.} Each row corresponds to an experimental condition, and the colors
correspond to the component number (ranked by proportion of variance
explained). The inflated brain plots display the top 20 highest-weighted hubs
(see \textit{Topographic Factor Analysis}) for each components'. The lists on
the right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
each components' brain map. Analogous maps computed separately for each story
segment may be found in Figure~\synthThirds.}

@@

@@

We measured properties of recorded neuroimaging data under different task
conditions that varied systematically in cognitive engagement and depth of
processing. We were especially interested in how \textit{informative} and
\textit{compressible} the activity patterns were under these different
conditions (Fig.~\ref{fig:information-compression}).

\subsection*{Functional neuroimaging data collected during story
  listening}

We examined an fMRI dataset collected by \cite{SimoEtal16} that the authors
have made publicly available at
\href{http://arks.princeton.edu/ark:/88435/dsp015d86p269k}{arks.princeton.edu/ark:/88435/dsp015d86p269k}.
The dataset comprises neuroimaging data collected as participants listened to
an audio recording of a story (intact condition; 36 participants), listened to
temporally scrambled recordings of the same story (17 participants in the
paragraph-scrambled condition listened to the paragraphs in a randomized order
and 36 in the word-scrambled condition listened to the words in a randomized
order), or lay resting with their eyes open in the scanner (rest condition; 36
participants). Full neuroimaging details may be found in the original paper for
which the data were collected~\cite{SimoEtal16}. Procedures were approved by
the Princeton University Committee on Activities Involving Human Subjects, and
by the Western Institutional Review Board (Puyallup, WA). All subjects were
native English speakers with normal hearing and provided written informed
consent.

\subsection*{Hierarchical topographic factor analysis (HTFA)}

Following our prior related work, we used HTFA~\cite{MannEtal18} to derive a
compact representation of the neuroimaging data. In brief, this approach
approximates the timeseries of voxel activations (44,415 voxels) using a much
smaller number of radial basis function (RBF) nodes~\citep[in this case, 700
nodes, as determined by an optimization procedure;][]{MannEtal18}). This
provides a convenient representation for examining full-brain activity patterns
and network dynamics. All of the analyses we carried out on the neuroimaging
dataset were performed in this lower-dimensional space. In other words, each
participant's data matrix, $\mathbf{X}$, was a number-of-timepoints ($T$) by 700
matrix of HTFA-derived factor weights (where the row and column labels were
matched across participants). Code for carrying out HTFA on fMRI data may be
found as part of the BrainIAK toolbox~\cite{CapoEtal17, KumaEtal21}, which may
be downloaded at \href{https://brainiak.org/}{brainiak.org}.

\subsection*{Principal components analysis (PCA)}

We applied group PCA~\citep{SmitEtal14} separately to the HTFA-derived
representations of the data (i.e., factor loadings) from each experimental
condition. Specifically, for each condition, we considered the set of all
participants' $T$ by 700 factor weight matrices. We used group PCA to project
these 700-dimensional matrices into a series of $k$-dimensional spaces, for $k
\in \{3, 4, 5, ..., 700\}$. This yielded a set of number-of-participants
matrices, each with $T$ rows and $k$ columns.

\subsection*{Temporal decoding}

We sought to identify neural patterns that reflected participants' ongoing
cognitive processing of incoming stimulus information. As reviewed by
\cite{SimoEtal16}, one way of homing in on these stimulus-driven neural
patterns is to compare activity patterns across individuals. In particular,
neural patterns will be similar across individuals to the extent that the
neural patterns under consideration are stimulus-driven, and to the extent that
the corresponding cognitive representations are reflected in similar spatial
patterns across people~\cite{SimoChan20}. Following this logic, we used an
across-participant temporal decoding test developed by \cite{MannEtal18} to
assess the degree to which different neural patterns reflected ongoing
stimulus-driven cognitive processing across people. The approach entails using
a subset of the data to train a classifier to decode stimulus timepoints (i.e.,
moments in the story participants listened to) from neural patterns. We use
decoding (forward inference) accuracy on held-out data, from held-out
participants, as a proxy for the extent to which the inputted neural patterns
reflected stimulus-driven cognitive processing in a similar way across
individuals.

\subsubsection*{Forward inference and decoding accuracy}

We used an across-participant correlation-based classifier to decode which
stimulus timepoint matched each timepoint's neural pattern. For a given value
of $k$ (i.e., number of principal components), we first used group PCA to
project the data from each condition into a $k$-dimensional space. Next, we
divided the participants into two groups: a template group,
$\mathcal{G}_{\mathrm{template}}$ (i.e., training data), and a to-be-decoded
group, $\mathcal{G}_{\mathrm{decode}}$ (i.e., test data). We averaged the
projected data within each group to obtain a single $T$ by $k$ matrix for each
group. Next, we correlated the rows of the two averaged matrices to form a $T$
by $T$ decoding matrix, $\mathbf{\Lambda}$. In this way, the rows of
$\mathbf{\Lambda}$ reflected timepoints from the template group, while the
columns reflected timepoints from the to-be-decoded group. We used
$\mathbf{\Lambda}$ to assign temporal labels to each timepoint (row) from the
test group's matrix, using the row of the training group's matrix with which it
was most highly correlated. We repeated this decoding procedure, but using
$\mathcal{G}_{\mathrm{decode}}$ as the template group and
$\mathcal{G}_{\mathrm{template}}$ as the to-be-decoded group. Given the true
timepoint labels (for each group), we defined the decoding accuracy as the
average proportion of correctly decoded timepoints, across both groups (where
chance perfomance is $\frac{1}{T}$). In Figures~\ref{fig:inflection}
and~\ref{fig:inflection-thirds} we report the decoding accuracy for each
condition and value of $k$, averaged across $n = 1,000$ cross validation folds.

\subsection*{Reverse inference}

To help interpret the brain activity patterns we found within the contexts of
other studies, we created summary maps of each principal component, for each
experimental condition, by summing together the 20 HTFA-derived RBF nodes (see
\textit{Hierarchical Topographic Factor Analysis}) with the highest absolute
value weights for each of the top 5 components (Figs.~\ref{fig:neurosynth-pca},
\synthThirds). We then carried out a meta analysis using
Neurosynth~\cite{RubiEtal17} to identify the 5 terms most commonly associated
with the given map.

\section*{Data and code availability}

All of the code used to produce the figures and results in this manuscript,
along with links to the corresponding datasets, may be found at 
\href{https://github.com/ContextLab/pca_paper}{github.com/ContextLab/pca\_paper}.

### 2023-02-27 paper/supplement.tex
finished draft of methods and supplemental materials

\newcommand{\synthThirds}{S1}

\title{High-order cognition is supported by complex but compressible brain activity patterns}

@@

\begin{abstract}

We applied dimensionality reduction algorithms and pattern classifiers to
functional neuroimaging data collected as participants listened to a story,
temporally scrambled versions of the story, or underwent a resting state
scanning session. These experimental conditions were intended to require
different depths of processing and inspire different levels of cognitive
engagement. We considered two primary aspects of the data. First, we treated
the number of features (components) required to achieve a threshold decoding
accuracy as a proxy for the ``compressibility'' of the neural patterns (where
fewer components indicate greater compression). Second, we treated the
maximum achievable decoding accuracy across participants as an indicator of the
``stability'' of the recorded patterns. Overall, we found that neural patterns
recorded as participants listened to the intact story required fewer features
to achieve comparable classification accuracy to the other experimental
conditions. However, the peak decoding accuracy (achievable with more features)
was also highest during intact story listening. Taken together, our work
suggests that our brain networks flexibly reconfigure according to ongoing task
demands, and that the activity patterns associated with higher-order cognition
and high engagement are both more complex and more compressible than the
activity patterns associated with lower-order tasks and lower levels of
engagement.

\end{abstract}

\doublespacing

\section*{Introduction}

Large-scale networks, including the human brain, may be conceptualized as
occupying one or more positions along on a continuum. At one extreme, every
node is fully independent of every other node. At the other extreme, all nodes
behave identically. Each extreme optimizes key properties of how the network
functions. When every node is independent, the network is maximally
\textit{expressive}: if we define the network's ``state'' as the activity
pattern across its nodes, then every state is equally reachable by a network
with fully independent nodes. On the other hand, a network of identically
behaved nodes optimizes \textit{robustness}: any subset of nodes may be removed
from the network without any loss of function or expressive power, as long as
any single node remains. Presumably, most natural systems tend to occupy
positions between these extremes. We wondered: might the human brain
reconfigure itself to be more flexible or more robust according to ongoing
demands? In other words, might the brain reconfigure its connections or
behaviors under different circumstances to change its position along this
continuum?

% compression (redundancy) vs. complexity -- appear to balance each other, butnot necessarily.

Closely related to the above notions of expressiveness versus robustness are
measures of how much \textit{information} is contained in a given signal or
pattern, and how \textit{redundant} a signal is~\citep{Shan48}. Formally,
information is defined as the amount of uncertainty about a given variables'
outcomes (i.e., entropy), measured in \textit{bits}, or the optimal number of
yes/no questions needed to reduce uncertainty about the variable's outcomes to
zero. Highly complex systems with many degrees of freedom (i.e., high
flexibility and expressiveness), are more information-rich than simpler or more
constrained systems. The redundancy of a signal denotes the difference how
expressive the signal \textit{could} be (i.e., proportional to the number of
unique states or symbols used to transmit the signal) and the actual
information rate (i.e., the entropy of each individual state or symbol). If a
brain network's nodes are fully independent, then the number of bits required
to express a single activity pattern is proportional to the number of nodes.
The network would also be minimally redundant, since the status of every node
would be needed to fully express a single brain activity pattern. If a brain
network's nodes are fully coupled and identical, then the number of bits
required to express a single activity pattern is proportional to the number of
unique states or values any individual node can take on. Such a network would
be highly redundant, since knowing any individual node's state would be
sufficient to recover the full-brain activity pattern. Highly redundant systems
are also robust, since there is little information loss from losing any given
observation.

We take as a given that brain activity is highly flexible: our brains can
exhibit nearly infinite activity patterns. This flexibility implies that our
brains activity patterns are highly information rich. However, brain activity
patterns are also highly structured. For example, full-brain correlation
matrices are stable within~\citep{FinnEtal15, FinnEtal17, GratEtal18} and
across~\citep{YeoEtal11, GlerEtal12, GratEtal18, ColeEtal14} individuals. This
stability suggests that our brains' activity patterns are at least partially
constrained, for example by anatomical, external, or internal factors.
Constraints on brain activity that limit its flexibility decrease
expressiveness (i.e., its information rate). However, constraints on brain
activity also increase its robustness to noise (e.g., ``missing'' or corrupted
signals may be partially recovered). For example, recent work has shown that
full-brain activity patterns may be reliably recovered from only a relatively
small number of implanted electrodes~\citep{OwenEtal20, ScanEtal21}. This
robustness property suggests that the relevant signal (e.g., underlying factors
that have some influence over brain activity patterns) are compressible.

\begin{figure}[tp]

  \centering
  \includegraphics[width=\textwidth]{figs/information_and_compressibility}

  \caption{\textbf{Information content and compressibility.} \textbf{A. Variance
  explained for two images.} We applied principal components analysis to a
  photograph and drawing, treating each row of the images as ``observations.''
  Across columns, we identified the number of components required to explain
  100\%, 95\%, 75\%, or 50\% of the cumulative variance in each image (the 100\%
  columns denote the original images). The numbers of components are indicate in
  parentheses, and the resulting ``compressed'' images are displayed. \textbf{B.
  Representing two images with different numbers of components.} Using the same
  principal component decompositions as in Panel A, we computed the cumulative
  proportion of variance explained with 380 (original images), 20, 4, or 2
  components. \textbf{C. Cumulative variance explained versus number of
  components.} For the images displayed in Panels A and B, we plot the cumulative
  proportion of variance explained as a function of the number of components used
  to represent each image. \textbf{D. Information rate and compressibility.}
  Across multiple images, the information rate (i.e., the amount of information
  contained in each image; horizontal axis) is high when each individual pixel
  provides information that cannot be inferred from other pixels.
  High-information rate images tend to be high-resolution, and low-information
  rate images tend to be low-resolution. Compressibility is related to the
  difference between the information required to specify the original versus
  compressed images (vertical axis). Highly compressible images often contain
  predictable structure (redundancies) that can be leveraged to represent the
  images much more efficiently than in their original feature spaces.}

  \label{fig:information-compression} 
  \end{figure}

To the extent that brain activity patterns contain rich task-relevant
information, we should be able to use the activity patterns to accurately
differentiate between different aspects of the task~\citep[e.g., using pattern
classifiers;][]{NormEtal06b}. For example, prior work has shown a direct
correspondence between classification accuracy and the information content of a
signal~\citep{Alva02}. To the extent that brain activity patterns are
compressible, we should be able to generate simplified (e.g., lower
dimensional) representations of the data while still preserving the relevant or
important aspects of the original signal. In general, information content and
compressibility are related but are partially dissociable
(Fig.~\ref{fig:information-compression}). If a given signal (e.g., a
representation of brain activity patterns) contains more information about
ongoing cognitive processes, then the peak decoding accuracy should be high.
And if the signal is compressible, a low-dimensional embedding of the signal
will be similarly informative to the original signal
(Fig.~\ref{fig:information-compression}D).

Several recent studies suggest that the complexity of brain activity is
task-dependent, whereby simpler tasks with lower cognitive demands are
reflected by simpler and more compressible brain activity patterns, and more
complex tasks with higher cognitive demands are reflected by more complex and
less compressible brain activity patterns~\citep{MackEtal20, OwenEtal21}. These
findings complement other work suggesting that functional connectivity
(correlation) patterns are task-dependent~\citep{FinnEtal17, OwenEtal20,
ColeEtal14}, although see~\cite{GratEtal18}. Higher-order cognitive processing
of a common stimulus also appears to drive more stereotyped task-related
activity and functional connectivity across individuals~\citep{HassEtal08,
LernEtal11, SimoChan20, SimoEtal16}.

The above prior studies are consistent with two potential descriptions of how
cognitive processes are reflects in brain activity patterns. One possibility is
that the information rate of brain activity increases during more complex or
higher-level cognitive processing. If so, then the ability to reliably decode
cognitive states from brain activity patterns should improve with task
complexity or with the level (or ``depth'') of cognitive processing. A second
possibility is that the compressibility of brain activity patterns increases
during more complex or higher-level cognitive processing. If so, then
individual features of brain recordings, or compressed representations of brain
recordings, should carry more information during complex or high-level (versus
simple or low-level) cognitive tasks.

We used a previously collected neuroimaging dataset to estimate the extent to
which each of these two possibilities might hold. The dataset we examined
comprised functional magnetic resonance imaging (fMRI) data collected as
participants listened to an audio recording of a 10-minute story, temporally
scrambled recordings of the story, or underwent a resting state
scan~\citep{SimoEtal16}. Each of these experimental conditions evokes different
depths of cognitive processing~\citep{SimoEtal16,LernEtal11,
HassEtal08,OwenEtal21}. We used across-participant classifiers to decode
listening times in each condition, as a proxy for how ``informative'' the
task-specific activity patterns were~\citep{SimoChan20}. We also use principle
components analysis to generate lower-dimensional representations of the
activity patterns. We then repeated the classification analyses after
preserving different numbers of components and examined how classification
accuracy changed across the different experimental conditions.

\section*{Results}

\begin{figure}
  \centering

  \includegraphics[width=0.8\textwidth]{figs/decoding_and_inflection}

\caption{\textbf{Decoding accuracy and compression.} \textbf{A. Decoding
accuracy by number of components.} Ribbons of each color display
cross-validated decoding performance for each condition (intact, paragraph,
word, and rest), as a function of the number of components (features) used to
represent the data. The horizontal red line denotes chance performance, and the
horizontal black line denotes 5\% decoding accuracy (used as a reference in
Panel B). \textbf{B. Numbers of components required to reach a fixed decoding
accuracy threshold, by condition.} The panel displays a zoomed-in view of the
inset in Panel A. Intersections between each condition's decoding accuracy
curve and the 5\% decoding accuracy reference line are marked by arrows.
\textbf{C. Estimating inflection points.} We sought to identify an ``inflection
point'' for each decoding curve, denoting the number of components at which the
decoding curve asymptotes. We fit sigmoid functions to each decoding curve
(left sub-panel) and then computed the minimum number of components where the
second derivative of the sigmoid was both positive and less than a threshold
value of 0.0001. \textbf{D. Inflection points by condition.} Each dot displays
the number of components ($x$-axis) and decoding accuracy ($y$-axis) at one
condition's inflection point. All error ribbons denote bootstrap-estimated 95\%
confidence intervals.}

    \label{fig:inflection}
  \end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figs/decode_pcs_network}

  \caption{\textbf{Network-specific decoding accuracy and compression.}
  \textbf{A. Decoding accuracy and number of components for network-specific
  inflection points.} We considered the seven networks identified by
  \cite{YeoEtal11}. We computed each network's inflection point, for each
  experimental condition, using the procedure described in
  Figure~\ref{fig:inflection}C. \textbf{B. Network-specific decoding
  accuracy.} Each of the seven networks are colored according to the decoding
  accuracy at the network's inflection point for the ``intact'' experimental
  condition (corresponding to the dark blue curve in Panel A). \textbf{C.
  Network-specific compression.} Each of the seven networks are colored
  according to the number of components at the network's inflection point for
  the intact experimental condition. Larger numbers of components reflect lower
  compressibility.}

  \label{fig:networks}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figs/information_and_compression_over_time}

\caption{\textbf{Changes in decoding accuracy and compression over time.}
\textbf{A. Decoding accuracy by number of components, by story segment.} Each
family of curves is plotted in the same format as Figure~\ref{fig:inflection}A
but reflects data only from one third of the dataset. \textbf{B. Inflection
points by condition and segment.} The dots re-plot the inflection points from
Figure~\ref{fig:inflection}D for reference. The numbers denote the inflection
points for each third of the dataset (1: first third; 2: second third; 3: third
third; colors denote experimental conditions). \textbf{C. Change in decoding
accuracy over time, by number of components.} For each number of components
($x$-axis) and condition (color), we fit a regression line to the decoding
accuracies obtained for the first, second, and third thirds of the dataset
(corresponding to the left, middle, and right columns of Panel A,
respectively). The $y$-axis denotes the slopes of the regression lines. All
error ribbons denote bootstrap-estimated 95\% confidence intervals.}

\label{fig:inflection-thirds}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/pca_neurosynth}

\caption{\textbf{Top terms associated with the highest-weighted components by
condition.} Each row corresponds to an experimental condition, and the colors
correspond to the component number (ranked by proportion of variance
explained). The inflated brain plots display the top 20 highest-weighted hubs
(see \textit{Topographic Factor Analysis}) for each components'. The lists on
the right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
each components' brain map. Analogous maps computed separately for each story
segment may be found in Figure~\synthThirds.}

\label{fig:neurosynth-pca}

\section*{Discussion}

\section*{Methods}

We measured properties of recorded neuroimaging data under different task
conditions that varied systematically in cognitive engagement and depth of
processing. We were especially interested in how \textit{informative} and
\textit{compressible} the activity patterns were under these different
conditions (Fig.~\ref{fig:information-compression}).

\subsection*{Functional neuroimaging data collected during story
  listening}

We examined an fMRI dataset collected by \cite{SimoEtal16} that the authors
have made publicly available at
\href{http://arks.princeton.edu/ark:/88435/dsp015d86p269k}{arks.princeton.edu/ark:/88435/dsp015d86p269k}.
The dataset comprises neuroimaging data collected as participants listened to
an audio recording of a story (intact condition; 36 participants), listened to
temporally scrambled recordings of the same story (17 participants in the
paragraph-scrambled condition listened to the paragraphs in a randomized order
and 36 in the word-scrambled condition listened to the words in a randomized
order), or lay resting with their eyes open in the scanner (rest condition; 36
participants). Full neuroimaging details may be found in the original paper for
which the data were collected~\cite{SimoEtal16}. Procedures were approved by
the Princeton University Committee on Activities Involving Human Subjects, and
by the Western Institutional Review Board (Puyallup, WA). All subjects were
native English speakers with normal hearing and provided written informed
consent.

\subsection*{Hierarchical topographic factor analysis (HTFA)}

Following our prior related work, we used HTFA~\cite{MannEtal18} to derive a
compact representation of the neuroimaging data. In brief, this approach
approximates the timeseries of voxel activations (44,415 voxels) using a much
smaller number of radial basis function (RBF) nodes~\citep[in this case, 700
nodes, as determined by an optimization procedure;][]{MannEtal18}). This
provides a convenient representation for examining full-brain activity patterns
and network dynamics. All of the analyses we carried out on the neuroimaging
dataset were performed in this lower-dimensional space. In other words, each
participant's data matrix, $\mathbf{X}$, was a number-of-timepoints ($T$) by 700
matrix of HTFA-derived factor weights (where the row and column labels were
matched across participants). Code for carrying out HTFA on fMRI data may be
found as part of the BrainIAK toolbox~\cite{CapoEtal17, KumaEtal21}, which may
be downloaded at \href{https://brainiak.org/}{brainiak.org}.

\subsection*{Principal components analysis (PCA)}

We applied group PCA~\citep{SmitEtal14} separately to the HTFA-derived
representations of the data (i.e., factor loadings) from each experimental
condition. Specifically, for each condition, we considered the set of all
participants' $T$ by 700 factor weight matrices. We used group PCA to project
these 700-dimensional matrices into a series of $k$-dimensional spaces, for $k
\in \{3, 4, 5, ..., 700\}$. This yielded a set of number-of-participants
matrices, each with $T$ rows and $k$ columns.

\subsection*{Temporal decoding}

We sought to identify neural patterns that reflected participants' ongoing
cognitive processing of incoming stimulus information. As reviewed by
\cite{SimoEtal16}, one way of homing in on these stimulus-driven neural
patterns is to compare activity patterns across individuals. In particular,
neural patterns will be similar across individuals to the extent that the
neural patterns under consideration are stimulus-driven, and to the extent that
the corresponding cognitive representations are reflected in similar spatial
patterns across people~\cite{SimoChan20}. Following this logic, we used an
across-participant temporal decoding test developed by \cite{MannEtal18} to
assess the degree to which different neural patterns reflected ongoing
stimulus-driven cognitive processing across people. The approach entails using
a subset of the data to train a classifier to decode stimulus timepoints (i.e.,
moments in the story participants listened to) from neural patterns. We use
decoding (forward inference) accuracy on held-out data, from held-out
participants, as a proxy for the extent to which the inputted neural patterns
reflected stimulus-driven cognitive processing in a similar way across
individuals.

\subsubsection*{Forward inference and decoding accuracy}

We used an across-participant correlation-based classifier to decode which
stimulus timepoint matched each timepoint's neural pattern. For a given value
of $k$ (i.e., number of principal components), we first used group PCA to
project the data from each condition into a $k$-dimensional space. Next, we
divided the participants into two groups: a template group,
$\mathcal{G}_{\mathrm{template}}$ (i.e., training data), and a to-be-decoded
group, $\mathcal{G}_{\mathrm{decode}}$ (i.e., test data). We averaged the
projected data within each group to obtain a single $T$ by $k$ matrix for each
group. Next, we correlated the rows of the two averaged matrices to form a $T$
by $T$ decoding matrix, $\mathbf{\Lambda}$. In this way, the rows of
$\mathbf{\Lambda}$ reflected timepoints from the template group, while the
columns reflected timepoints from the to-be-decoded group. We used
$\mathbf{\Lambda}$ to assign temporal labels to each timepoint (row) from the
test group's matrix, using the row of the training group's matrix with which it
was most highly correlated. We repeated this decoding procedure, but using
$\mathcal{G}_{\mathrm{decode}}$ as the template group and
$\mathcal{G}_{\mathrm{template}}$ as the to-be-decoded group. Given the true
timepoint labels (for each group), we defined the decoding accuracy as the
average proportion of correctly decoded timepoints, across both groups (where
chance perfomance is $\frac{1}{T}$). In Figures~\ref{fig:inflection}
and~\ref{fig:inflection-thirds} we report the decoding accuracy for each
condition and value of $k$, averaged across $n = 1,000$ cross validation folds.

\subsection*{Reverse inference}

To help interpret the brain activity patterns we found within the contexts of
other studies, we created summary maps of each principal component, for each
experimental condition, by summing together the 20 HTFA-derived RBF nodes (see
\textit{Hierarchical Topographic Factor Analysis}) with the highest absolute
value weights for each of the top 5 components (Figs.~\ref{fig:neurosynth-pca},
\synthThirds). We then carried out a meta analysis using
Neurosynth~\cite{RubiEtal17} to identify the 5 terms most commonly associated
with the given map.

\section*{Data and code availability}

All of the code used to produce the figures and results in this manuscript,
along with links to the corresponding datasets, may be found at 
\href{https://github.com/ContextLab/pca_paper}{github.com/ContextLab/pca\_paper}.

\section*{Acknowledgements}

We acknowledge discussions with Rick Betzel, Emily Finn, and Jim Haxby. Our
work was supported in part by NSF CAREER Award Number 2145172 to J.R.M. The
content is solely the responsibility of the authors and does not necessarily
represent the official views of our supporting organizations. The funders had
no role in study design, data collection and analysis, decision to publish, or
preparation of the manuscript.

\section*{Author contributions}

Conceptulization: J.R.M. and L.L.W.O. Methodology: J.R.M. and L.L.W.O.
Implementation: L.L.W.O. Analysis: L.L.W.O. Writing, Reviewing, and Editing:
J.R.M. and L.L.W.O. Supervision: J.R.M.

\bibliographystyle{apacite}
\bibliography{CDL-bibliography/cdl}

>>>>

\newcommand{\synth}{1}
\renewcommand{\thefigure}{S\arabic{figure}}

\title{\textit{Supplemental materials for:} High-order cognition is supported by complex but compressible brain activity patterns}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/pca_neurosynth_thirds}

\caption{\textbf{Top terms associated with the highest-weighted components by
condition, broken down by story segment.} Each group of three rows corresponds
to an experimental condition, and the colors correspond to the component number
(ranked by proportion of variance explained). The rows' numbering denotes the
story segment used to compute each map (1: first third; 2: second third; 3:
third third). The inflated brain plots display the top 20 highest-weighted hubs
(see \textit{Topographic Factor Analysis}) for each components. The lists on
the right display the top five Neurosynth terms~\citep{RubiEtal17} decoded from
each components' brain map. Analogous maps computed for the entire story may be
found in Figure~\synth.}

\label{fig:neurosynth-thirds}

@@

\newpage
\renewcommand{\refname}{Supplemental references}
\bibliographystyle{apacite}
\bibliography{CDL-bibliography/cdl}
